Exploiting the Social Graph: Increasing Engagement in Collaborative Interactive Evolution

### 2017-01-27 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Spell checking

\section{Introduction}
Interactive Evolutionary Algorithms (IEAs) are, in general, standard EAs whose
fitness evaluations are performed by persons within an interactive 
system \cite{eiben2015interactive}.  Thus, the main evolutionary loop needs the intervention of a
human to perform the quality assessment of each proposed solution.
These kind of IEAs have demonstrated their ability for effectively
producing art and design \cite{Bentley:1999:intro,Sims:1991,todd:1992},
as well as other products in many other domains \cite{ie1}, and this
success has led to the development of web based systems such as \cite{picbreeder},
who depend on volunteers who visit the system to interactively help
with the search, using both anonymous and registered users. This
volunteeer system lowers the requirements of anyone participating in
the experiment thus increasing the {\em performance}, in terms of {\em
  human brain cycles}, of the whole system.

@@

Collaborative IEC (C-IEC) systems also need to store highly connected data, as it is common in current applications
like social networks. In order to deal with large datasets of connected data found in these systems,
graph databases \cite{angles2012comparison} have been proposed as an alternative to relational databases 
which have performance limitations when dealing with highly connected data \cite{holzschuher2013performance}.
In this work the use of a graph data model for representing the interactive process of C-IEC systems is proposed.
Moreover, an proof-of-concept implemantation of an IEA incorporating the graph model is presented.

@@

Before launch, each deployment was first tried with a few beta testers. When applying the leader board
gammification technique for the first time a problem was found in this stage. The problem was that some 
users were chating by giving a rating to an animation even before it was returned from the server, this was done by just
constantly clicking the mouse button. This is a common problem found in systems using leader boards because
by making the scores visible to other players they are encouraged to compete \cite{hickman2010total}. 
According to \cite{kumar2013gamification} best-practice gamification designs try to
refrain from using this element because of  problematic consequences of competition, which can result 
in negative conduct, low cooperation and collaboration, or disadvantaging certain player
demographics such as women. In order to still measure the effect of the leader board on the system, the 
problem of cheating was mitigated by disabling the rate button for a certain amount of time giving the animation
enough time to finish. But the problem persists, competitive users will try to get in to the leader board by other means.

% Future work could compare the use of gamification with out leaderboards
%
The results of each of the three experiments in terms of participation are detailed next.

\subsubsection{ Base deployment}
After a week of the anouncement the total number of volunteers for this experiment was 53 registered users. 
The total number of nodes generated in the graph was 595, with a total of 2220 relationships. In this
experiment a total of 500 new phenotypes where generated. The total number of evaluated
phenotypes for each volunteer is presented in figure
\ref{fig:top-ranked-participation} with a blue line. % Ranked by... - JJ

@@

When monitoring the experiments, using the graph database and the Cypher query language 
provided interesting views on the connected data difficult to replicate with a relational database. 
As an example we can consider comparing the difference between the number of volunteers in deployments
B and GG (shown in figures \ref{fig:B-network} and \ref{fig:GG-network} respectivly) additional
metrics and patterns are available when they are treated as a
network. %Please read the whole sentence and maybe rewrite - JJ
The same can be applied to 
the population, or the interaction as a whole. The graph could be used also by the evolutionary
algorithm, for instance when creating a new phenotype 
parents could be selected only if there is a minimum distance between them in the graph.

>>>>

\section{Introduction}
Interactive Evolutionary Algorithms (IEAs) are, in general, standard EAs whose
fitness evaluations are performed by persons within an interactive 
system \cite{eiben2015interactive}.  Thus, the main evolutionary loop needs the intervention of a
human to perform the quality assessment of each proposed solution.
These kind of IEAs have demonstrated their ability for effectively
producing art and design \cite{Bentley:1999:intro,Sims:1991,todd:1992},
as well as other products in many other domains \cite{ie1}, and this
success has led to the development of web based systems such as \cite{picbreeder},
who depend on volunteers who visit the system to interactively help
with the search, using both anonymous and registered users. This
volunteer system lowers the requirements of anyone participating in
the experiment thus increasing the {\em performance}, in terms of {\em
  human brain cycles}, of the whole system.

@@

Collaborative IEC (C-IEC) systems also need to store highly connected data, as it is common in current applications
like social networks. In order to deal with large datasets of connected data found in these systems,
graph databases \cite{angles2012comparison} have been proposed as an alternative to relational databases 
which have performance limitations when dealing with highly connected data \cite{holzschuher2013performance}.
In this work the use of a graph data model for representing the interactive process of C-IEC systems is proposed.
Moreover, an proof-of-concept implementation of an IEA incorporating the graph model is presented.

@@

Before launch, each deployment was first tried with a few beta testers. When applying the leader board
gamification technique for the first time a problem was found in this stage. The problem was that some 
users were cheating by giving a rating to an animation even before it was returned from the server, this was done by just
constantly clicking the mouse button. This is a common problem found in systems using leader boards because
by making the scores visible to other players they are encouraged to compete \cite{hickman2010total}. 
According to \cite{kumar2013gamification} best-practice gamification designs try to
refrain from using this element because of  problematic consequences of competition, which can result 
in negative conduct, low cooperation and collaboration, or disadvantaging certain player
demographics such as women. In order to still measure the effect of the leader board on the system, the 
problem of cheating was mitigated by disabling the rate button for a certain amount of time giving the animation
enough time to finish. But the problem persists, competitive users will try to get in to the leader board by other means.

The results of each of the three experiments in terms of participation are detailed next.

\subsubsection{ Base deployment}
After a week of the announcement the total number of volunteers for this experiment was 53 registered users. 
The total number of nodes generated in the graph was 595, with a total of 2220 relationships. In this
experiment a total of 500 new phenotypes where generated. The total number of evaluated
phenotypes for each volunteer is presented in figure
\ref{fig:top-ranked-participation} with a blue line. % Ranked by... - JJ

@@

When monitoring the experiments, using the graph database and the Cypher query language 
provided interesting views on the connected data difficult to replicate with a relational database. 
As an example we can consider comparing the difference between the number of volunteers in deployments
B and GG (shown in figures \ref{fig:B-network} and \ref{fig:GG-network} respectively) additional
metrics and patterns are available when they are treated as a
network. %Please read the whole sentence and maybe rewrite - JJ
The same can be applied to 
the population, or the interaction as a whole. The graph could be used also by the evolutionary
algorithm, for instance when creating a new phenotype 
parents could be selected only if there is a minimum distance between them in the graph.
